Interactively Visualizing Multivariate Market Segmentation Using the R Package Lionfish

### 2024-10-03 manuscript.tex
lots of minor tweaks

The \texttt{lionfish} GUI can be launched using the function \texttt{interactive\textunderscore tour()}. At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots. The dataset must be supplied as a numeric matrix, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}. The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour. The \texttt{obj} element further defines the properties of the chosen display. For example, to create a 2-dimensional tour, the user must provide a \texttt{tour\textunderscore history} object, which can be generated using the \texttt{tourr} package. For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed. The user can optionally specify the feature names, the arrangement of the plots, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.

@@

The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects. Below this, there are three additional interfaces (Figure \ref{fig:GUI_overview}F). The first can be used to adjust the number of bins in histograms—more bins result in higher resolution but slower display updates. The second allows users to animate tours, automatically shifting to the next projection after a specified amount of time. The third offers the ability to hide projection axes with a norm smaller than a chosen threshold, helping to reduce clutter. Users can also save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}G). New tours can be started using the current settings via the interface at the bottom (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g. heatmaps (Figure \ref{fig:GUI_overview}I).

@@

One can move through the projections of a tour by pressing the arrow keys or by specifying the index of the projection the be displayed and pressing the \texttt{Update frames} button. Users can animate the tours by toggling the \texttt{Animate} checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.

@@

More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.

@@

The majority of \texttt{lionfish} was written in \proglang{Python} \citep{python}, while the \proglang{R} \citep{R} side of the package handles setting up the \proglang{Python} environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface. To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from \proglang{R} to \proglang{Python}, the \texttt{reticulate} package was used. This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr}.

@@

Before launching the \texttt{lionfish} interface the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface. We launch the interface to explore and potentially refine this solution.

@@

The last, $f^f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature. It is useful to examine how features are related to a cluster. Considering the last metric, for example, if $f^f_{i1} = (0.2, 0.7, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in cluster 2. In the context of the later examples, this would mean activity 1 is most commonly listed in cluster 2. This normalisation produces what is called the {\bf intra-feature fraction}. It also has to be noted that this metric is heavily influenced by the cluster sizes. Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction.

@@

This process allows us to visually assess the separation and similarities between the clusters, providing insight into the structure of the dataset. By highlighting each cluster individually, we can evaluate their distinctiveness in different projections. The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, and going to the spa. The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar in the selected projection.

@@

\subsubsection{Segmentation of travellers travelling without friends}

In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling without friends. We can assume that this group consits of tourists travelling alone as couples or with their family. Additional features to further subsegment these groups might be interesting for future surveys. As a tourist agency, we might be interested in targeting these travellers more effectively. However, subsets 2 and 6 also include individuals who enjoy spending time with their friends. To further explore the dataset, we can launch an \texttt{interactive\_tour()} with the same configuration as in the previous example. To achieve better separation between the subsets, we can skip to the last frames of a 2 dimensional tour optimized for the LDA index.

@@

We also observe distinctive differences in the interests of the subsets that traveled without friends. Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms. Given that Subset 7 showed a notable interest in going to pubs, we can assume that the solo travelers and couples in this subset are looking to meet new people. This insight could be leveraged in a marketing campaign by bundling activities that appeal to Subset 7, creating packages tailored to these tourists. Such packages would provide opportunities for them to connect with other travelers who share similar interests while enjoying their preferred activities. Additionally, some solo travelers and couples might prefer experiences or environments without children, allowing for the creation of adult-oriented packages. The remaining subgroup within Subset 7 consists of families. To better cater to this group, packages can be refined to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.

@@

Future developments in the software might include expanding the range of interactive features and providing additional visualization methods. Because it is built on \proglang{R} and \proglang{Python}, it would also be possible to integrate more closely with the machine learning algorithms, such as generating new cluster solutions directly from the GUI. For more complex shapes, containing concavities, or non-linear boundaries, implementation of sliced tours~\citep{Laa2020}, would be recommended. The integration of new plot types could also be streamlined to improve the versitility of the package.

Furthermore, \texttt{lionfish} provides an additional demonstration of how \proglang{R} and \proglang{Python} can be integrated to harness the strengths of both programming languages.

>>>>

The \texttt{lionfish} GUI can be launched using the function \texttt{interactive\textunderscore tour()}. At a minimum, the user needs to provide both the dataset and the instructions for constructing the desired plots. The dataset must be supplied as a numeric matrix, while the plotting instructions should be passed as a list containing the named elements \texttt{type} and \texttt{obj}. The \texttt{type} element specifies the type of display to generate, such as \texttt{scatter} for a scatterplot or \texttt{2d\textunderscore tour} for a 2-dimensional tour. The \texttt{obj} element further defines the properties of the chosen display. For example, to create a 2-dimensional tour, the user must provide a \texttt{tour\textunderscore history} object, which can be generated using the \texttt{tourr} package. For a scatterplot, the user needs to provide a vector of strings specifying the names of the features to be displayed. With the optional arguments of \texttt{interactive\textunderscore tour()} the user can specify the feature names, the arrangement of the plots within the GUI, predefined subsets of the data (e.g., cluster solutions), custom names for these subsets, the number of available subsets, and the size of each plot.

@@

The frame selection interface (Figure \ref{fig:GUI_overview}E) displays the frames of the currently shown tours and enables users to jump between frames of the tour objects. Below this, there are three additional interfaces (Figure \ref{fig:GUI_overview}F). The first can be used to adjust the number of bins in histograms—more bins result in higher resolution but slower display updates. The second allows users to animate tours, automatically shifting to the next projection after a specified amount of time. The third offers the ability to hide projection axes with a norm smaller than a chosen threshold, helping to reduce clutter. Users can also save and load projections and subsets using the respective buttons (Figure \ref{fig:GUI_overview}G). New tours can also be started directly from within the GUI (Figure \ref{fig:GUI_overview}H), and users can select different metrics for some plot types e.g. heatmaps (Figure \ref{fig:GUI_overview}I).

@@

One can move through the projections of a tour by pressing the arrow keys or by specifying the frame of the projection to be displayed and pressing the \texttt{Update frames} button. Users can animate the tours by toggling the \texttt{Animate} checkbox and specifying a time interval, so that the displays automatically move to the next projections after the specified time interval.

@@

More information on all the features of the GUI can be found on \href{https://github.com/mmedl94/lionfish}{https://github.com/mmedl94/lionfish} and on \href{https://mmedl94.github.io/lionfish/}{https://mmedl94.github.io/lionfish/}.

@@

The majority of \texttt{lionfish} was written in \proglang{Python} \citep{python}. The \proglang{R} \citep{R} side of the package handles setting up the \proglang{Python} environment within the interactive interface is being run, launching the interactive tour, and generating new tours when initiated through the interface. To incorporate the functionality of the \texttt{tourr} package without translating large portions of its code from \proglang{R} to \proglang{Python}, the \texttt{reticulate} package was used. This approach allows \texttt{lionfish} to automatically benefit from updates to \texttt{tourr} and avoided the necessity to reimplement code in \proglang{Python} that was already written in \proglang{R}.

@@

Before launching the \texttt{lionfish} interface, the user should have performed a partitioning of their choice, and provide the initial clustering solution used by the interface. We launch the interface to explore and potentially refine this solution.

@@

The last, $f^f$ is normalized relative to each feature, which can be considered to be the distribution of clusters on each feature. It is useful to examine how features are related to a cluster. Considering the last metric, for example, if $f^f_{i1} = (0.2, 0.7, 0, 0.1, 0)$ ($k=5$) would indicate that high values on feature 1 primarily are in cluster 2. In the context of the later examples, this would mean activity 1 is most commonly listed in cluster 2. This normalisation produces what is called the {\bf intra-feature fraction}. It also has to be noted that this metric is heavily influenced by the cluster sizes. Consider a feature that is uniformly distributed across all clusters, then the largest cluster would receive the highest intra-feature fraction for that feature.

@@

This process allows us to visually assess the separation and similarities between the clusters, providing insight into the structure of the dataset. By highlighting each cluster individually, we can evaluate their distinctiveness in different projections. The most influential features shown in Figure \ref{fig:winter_activ_cluster_highlights} are pool/sauna, alpine skiing, museums, going to the spa, going for walks and sight seeing. The projection roughly separates clusters 1 (blue), 2 (orange), and 3 (green) from each other and the other three clusters (red, violet and brown), which appear to be quite similar in the selected projection.

@@

\subsubsection{Segmentation of tourists travelling without friends}

In the heatmap displaying the intra-cluster fraction shown in Figure \ref{fig:aus_heatmapg}, we observe that subsets 1, 2, and 6 tend to prefer travelling without friends. We can assume that this group consists of tourists travelling alone as couples or with their family. Additional features to further subsegment these groups might be interesting for future surveys. As a tourist agency, we might be interested in targeting these travellers more effectively. However, subsets 2 and 6 also include individuals who enjoy spending time with their friends. To further explore the dataset, we can launch an \texttt{interactive\_tour()} with the same configuration as in the previous example. To achieve better separation between the subsets, we can skip to the last frames of a 2 dimensional tour optimized for the LDA index.

@@

We also observe distinctive differences in the interests of the subsets that traveled without friends. Subset 7 was interested in relaxing, shopping, sightseeing, wildlife, going to the beach, visiting pubs, and exploring farms. Given that Subset 7 showed a notable interest in going to pubs, we can assume that the solo travelers and couples in this subset are looking to meet new people. This insight could be leveraged in a marketing campaign by bundling activities that appeal to Subset 7, creating packages tailored to these tourists. Such packages would provide opportunities for them to connect with other travelers who share similar interests while enjoying their preferred activities. Additionally, some solo travelers and couples might prefer experiences or environments without children, allowing for the creation of adult-oriented packages. The remaining subgroup within Subset 7 consists of families. To better cater to this group, packages can be refined for instance to consider that not all museums or festivals are equally suitable for children, enabling more family-friendly customizations.

@@

Future developments in \texttt{lionfish} might include expanding the range of visualization methods and offering additional interactive features, such as enabling the generation of new cluster solutions directly from the GUI. Leveraging the integration of both \proglang{R} and \proglang{Python}, future enhancements could include a seamless combination of algorithms from each language, which broadens the expansion potential of the software. For more complex data shapes, such as those with concavities or non-linear boundaries, the implementation of sliced tours~\citep{Laa2020} is recommended to improve exploratory analysis. Streamlining the addition of new plot types could also further enhance the versatility of the package, making it more adaptable for different data visualization needs. This ongoing integration of \proglang{R} and \proglang{Python} exemplifies how the strengths of both languages can be harnessed for the development of new software packages.
